---
title: "Homework 6"
author: "Xinyu Yang"
date: "11/22/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(modelr)
library(mgcv)
```

## Problem 1

```{r}
birthweight <-
  read_csv("./data/birthweight.csv") %>% 
  janitor::clean_names() %>% 
  filter(frace != 9) %>% 
  select(-pnumlbw,-pnumsga) %>% 
  mutate(
    babysex = as.factor(babysex),
    frace = as.factor(frace),
    mrace = as.factor(mrace),
    malform = as.factor(malform)
  )

#check missing data
apply(is.na(birthweight),2,sum)
```

After read in the dataset, no missing value was identified; variables babysex, frace, mrace, malform were converted into factor; vairalbes pnumlbw, pnumgsa contained only value of 0, thus were excluded; frace level 9 (unknown) was also excluded from the analysis.

```{r}
#function for bivariate regression
bivariate = function(x) {
  lm(bwt ~ x, data = birthweight) %>% broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  filter(term != "(Intercept)") %>% 
  mutate(p.value = round(p.value, 4))
}

bivariate_result <-
  birthweight %>% 
  map_dfr(~bivariate(.), .id = "source")

#fit model1 with selected vairables
model1 <- lm(bwt ~ babysex + gaweeks + momage + mrace + ppbmi + wtgain, data = birthweight) 
model1 %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  mutate(
    p.value = round(p.value, 3),
    estimate = round(estimate, 2),
    ) %>% 
  knitr::kable()
```

Describe modeling process...


Plot residuals against fitted value
```{r}
birthweight %>% 
modelr::add_residuals(model1) %>% 
modelr::add_predictions(model1) %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = 0.3)

```

The residual vs. predicted value is relatively shapeless without clear patterns in the data, and it is generally symmetrically distributed around the 0 line.

```{r}
#model2: length at birth and gestational age as predictors (main effects only)
model2 <- lm(bwt ~ blength + gaweeks, data = birthweight) 

model2 %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  mutate(
    p.value = round(p.value, 3),
    estimate = round(estimate, 2),
    ) %>% 
  knitr::kable()

#model3: head circumference, length, sex, and all interactions (including the three-way interaction)
model3 <- lm(bwt ~ (bhead + blength + babysex)^3, data = birthweight)

model3 %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  mutate(
    p.value = round(p.value, 3),
    estimate = round(estimate, 2),
    ) %>% 
  knitr::kable()
```

Cross validation to compare models:
```{r}
cv_df = 
  crossv_mc(birthweight, 100)

cv_df =
  cv_df %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))

cv_df = 
  cv_df %>% 
  mutate(
    mod_1 = map(train, ~lm(bwt ~ babysex + gaweeks + momage + mrace + ppbmi + wtgain, data = .x)),
    mod_2 = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
    mod_3 = map(train, ~lm(bwt ~ (bhead + blength + babysex)^3, data = .x))
    ) %>% 
  mutate(
    rmse_1 = map2_dbl(mod_1, test, ~rmse(model = .x, data = .y)),
    rmse_2 = map2_dbl(mod_2, test, ~rmse(model = .x, data = .y)),
    rmse_3 = map2_dbl(mod_3, test, ~rmse(model = .x, data = .y)))

cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin() +
  theme_minimal()
```

From the violin plot we can see that the prediction error is the lowest with model 3 (using head circumference, length, sex, and all interactions).

## Problem 2
```{r message=FALSE}
#read in dataset
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

temperature <-
weather_df %>% 
  select(tmax, tmin)

```

Bootstrap samples:
```{r}
boot_sample = function(df) {
  sample_frac(df, replace = TRUE)
}

boot_straps = 
  data_frame(
    strap_number = 1:5000,
    strap_sample = rerun(5000, boot_sample(temperature))
  )

bootstrap_results = 
  boot_straps %>% 
  mutate(
    models = map(strap_sample, ~lm(tmax ~ tmin, data = .x) ),
    results = map(models, broom::tidy),
    overall = map(models, broom::glance)) 

#log(beta0*beta1) and its 95% CI
bootstrap_results %>% 
  select(-strap_sample, -models, -overall) %>% 
  unnest(cols = c(results)) %>% 
  select(strap_number, term, estimate) %>% 
  pivot_wider(names_from = term, values_from = estimate) %>% 
  rename(beta0 = "(Intercept)", beta1 = tmin) %>% 
  mutate(log_beta = log((beta0*beta1))) %>% 
  summarise(log_95 = list(enframe(quantile(log_beta, probs = c(0.025, 0.5, 0.975))))) %>% 
  unnest() %>% 
  rename(percentile = name) %>% 
  knitr::kable()
```

The 95% confidence interval for log(beta0*beta1) is 

```{r}
#r-squared and 95% CI
bootstrap_results %>% 
  select(-strap_sample, -models, -results) %>% 
  unnest(cols = c(overall)) %>% 
  janitor::clean_names() %>% 
  select(r_squared) %>% 
  summarise(r_95 = list(enframe(quantile(r_squared, probs = c(0.025, 0.5, 0.975))))) %>% 
  unnest() %>% 
  rename(percentile = name) %>% 
  knitr::kable()
```

